---
title: "Tutorial 2 Answers"
author: "Ryan Cordell, Lindsay Thomas"
date: "2/25/2019"
output: html_document
---
# Answers From Tutorial

```{r}
# print the first lines of the census dataframe

head(census)

# print the last lines of the census dataframe

tail(census)

# print the first or last lines of one specific column in the census dataframe

head(census$TotalPopulation)

```


```{r}
# import the same census data, using a new variable

census2 <- read.csv(file = "/Users/lxt308/Dropbox/Teaching/2019-spring/eng-682/tutorials/data/1840-census-data.csv")

# select 3-4 columns of interest

census2 <- dplyr::select(census2, QualifyingAreaName, TotalPopulation, Male, Female)

# filter the data frame by a value in one of those columns

census_highpop <- filter(census2, TotalPopulation >= 2000)
```


# Exercises

1. Create a new column that shows the ratio of `FreeColoredPopulation` to `SlavePopulation` in each county. You may need to repeat some of the import steps above, and you can use a new variable if you wish. 

```{r}
# write your code here
census_pop <- read.csv(file = "/Users/lxt308/Dropbox/Teaching/2019-spring/eng-682/tutorials/data/1840-census-data.csv")
census_pop <- dplyr::rename(census_pop, county = QualifyingAreaName)
census_pop <- separate(census_pop, county, into = c("county", "state"), sep = "\\, ")
census_pop <- select(census_pop, county, state, FreeColoredPopulation, SlavePopulation)
census_pop <- group_by(census_pop, county)
census_pop <- mutate(census_pop, ratio = (FreeColoredPopulation/SlavePopulation)*100)
```

2. Create a "long" dataframe with columns for "publication_type" (e.g. "Newspaper") and "count". 

```{r}
census <- read.csv(file="/Users/lxt308/Dropbox/Teaching/2019-spring/eng-682/tutorials/data/1840-census-data.csv")[ , c("QualifyingAreaName", "Newspapers", "Newspapers_Daily", "Newspapers_Weekly", "Newspapers_SemiTriWeekly", "Periodicals", "PrintingOffices")]
census <- rename(census, county = QualifyingAreaName)
census_long <- gather(census, "publication_type", "count", 2:7)
census_long <- separate(census_long, county, into = c("county", "state"), sep = "\\, ")
census_long <- na.omit(census_long)
```

3. Determine the state (not the county!) with the highest population of literate whites over 20 years old, then determine the state with the highest ratio of literate whites to general population.

```{r}
census <- read_csv(file="/Users/lxt308/Dropbox/Teaching/2019-spring/eng-682/tutorials/data/1840-census-data.csv")
census <- rename(census, county = QualifyingAreaName)
census <- separate(census, county, into = c("county", "state"), sep = "\\, ")
census <- group_by(census, state)
literate <- summarize(census, totalLit = sum(LiterateWhiteAge20andOver))
literate <- arrange(literate, desc(totalLit))

```

```{r}
census <- read_csv(file="/Users/lxt308/Dropbox/Teaching/2019-spring/eng-682/tutorials/data/1840-census-data.csv")
census <- rename(census, county = QualifyingAreaName)
census <- separate(census, county, into = c("county", "state"), sep = "\\, ")
census <- group_by(census, state)
literate <- summarize(census, totalLit = sum(LiterateWhiteAge20andOver) / sum(TotalPopulation))
literate <- arrange(literate, desc(totalLit))

```