---
title: "Tutorial 3 Answers"
author: "Ryan Cordell, Jonathan Fitzgerald, Lindsay Thomas"
date: "3/4/2019"
output: html_document
---
# Answers From Tutorial

```{r}
# inspect the top rows, or should we say the *head*, of the wordcounts data frame 
head(wordcounts)

```


# Exercises

1. Above, we created a concordance for the single text. Can you create a concordance for the entire set of texts from the Wright corpus? BONUS: You'll probably notice that your new data frame's columns seem a bit out of order, can you rearrange them so that they are: filename, word, word2, word3? (Hint: Check out the `select()` function in the `dplyr` package.)

```{r}
allWRIGHTwords %>% mutate(word2 = lead(word,1)) %>% head
multiColumn2 <- allWRIGHTwords %>% mutate(word2 = lead(word,1),word3=lead(word,2),word4=lead(word,3)) %>% select(filename, word, word2, word3, word4)

```

Alternatively, you can rearrange the columns like this:

```{r}

multiColumn2 = allWRIGHTwords[,c("filename","word","word2","word3")]

```
